\MakeUppercase{The Structure of $k$-Planar Graphs}

### 2019-12-04 k-planar.tex
Fixed mistake caused by k+1 versus k

\begin{thm}
\label{kPlanarBasic}
Every $k$-planar graph is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth $(18k^2+30k)\binom{k+4}{3}-1$ and for some path $P$. 
\end{thm}

@@

\begin{thm}
\label{k-planar}
Every $k$-planar graph has an $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\binom{k+4}{3}-1$. %$\treewidth$.
\end{thm}

@@

\begin{thm}
\label{g-k-planar}
Every $(g,k)$-planar graph has an $H$-partition of layered width at most 
$\max\{2g,3\}\cdot(6k^2 + 10k)$ in which $H$ has treewidth at most $\binom{k+5}{4}-1$.
\end{thm}

@@

We can reduce the layered width of the $H$-partition of $G$ from $O(k^3)$ to $O(k^2)$ by observing that the dummy vertices in $V(G_0)\setminus V(G)$ do not contribute to the layered width of this partition.  In this setting, the proof of \cref{clm:general-width} is simpler since each vertex $w\in Y_x$ contributes at most two vertices to $V_i\cap Y_x$.  More precisely, each path $P\in\mathcal{P}$ containing an internal (dummy) vertex $w\in Y_x\cap (V_{i-j}\cup V_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap V_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap V_j$ for $j\in\{\floor{(k+1)/2}+1,\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap V_j$ for $j > k+1$.
Redoing the calculation at the end of the proof of \cref{clm:general-width} then yields 
\begin{align*}
    |S_x\cap Y_i| \le d\ell\left( 
       2 
       + 4\left\lfloor\frac{k+1}{2}\right\rfloor 
       + 2\left\lceil\frac{k+1}{2}\right\rceil
       \right)
       & =
       d\ell\left(
         2 + 2(k+1) + 2\left\lfloor\frac{k+1}{2}\right\rfloor
       \right) \\
       & \le
       d\ell(3k+5) 
      = 18k+30 \enspace .
\end{align*}    
With this change, the layered width of the partition given by \cref{modern-major-general} becomes $18k^2+30k$.  This establishes \cref{k-planar}.

@@

\begin{cor}\corlabel{first}
  For every $k$-planar graph $G$, 
  %$\pi(G)\le 12(2k+3)(k+1)\times 4^{(k+2)(k+3)(k+4)/6}$.
  $\pi(G)\le (18k^2 + 30k) 4^{\binom{k+4}{3}}$.
%  Every $k$-planar graph has a layered $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\binom{k+4}{3}}-1$.
\end{cor}

@@

\begin{cor}
  For every $(g,k)$-planar graph $G$, 
  \[ 
%    \pi(G)\le \max\{4g(3k+1),12(2k+3)(k+1)\}\times 4^{(k+2)(k+3)(k+4)/6} 
    \pi(G)\le \max\{2g,3\}\cdot(6k^2 + 10k) 4^{\binom{k+5}{4}}. 
     \]
%  \label{g-k-planar} Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most  
% $\max\{2g,3\}\cdot(6k^2 + 10k)$ in which $H$ has treewidth at most $\binom{k+5}{4}$.
\end{cor}

@@

\begin{cor}
\label{pck}
For every $k$-planar graph $G$, 
%for some integer $t\leq \binom{k+4}{3}-1$, 
%$$\chi_p(G)\le (18k^2+30k)(p+1) (t+1)\tbinom{p+t}{t}\enspace.$$
\[ \chi_p(G)\le (18k^2+30k)(p+1) \binom{k+4}{3} \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}\enspace. \]
%\note{DW}{I previously wrote ``for some integer $t\leq \binom{k+4}{3}-1$'' thinking that $\binom{p+t}{t}$ might not be monotone with $t$, but it is because of the $+t$.}
%Every $k$-planar graph has a layered $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\binom{k+4}{3}-1$. %$\treewidth$.
%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$. 
\end{cor}

@@

\begin{cor}
\label{pcgk}
For every $(g,k)$-planar graph $G$,  
%for some integer $t\leq \binom{k+5}{4}-1$, 
%$$\chi_p(G) \le \max\{2g,3\}\cdot(6k^2 + 10k) (p+1) (t+1)\tbinom{p+t}{t}\enspace.$$
\[ \chi_p(G) \le \max\{2g,3\}\cdot(6k^2 + 10k) (p+1) \binom{k+5}{4}\binom{p+\binom{k+5}{4}-1}{\binom{k+5}{4}-1}\enspace. \]
%O(gk^2)\times p^{O(k^3)}$. 
%%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$. 
%Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most 
%$\max\{2g,3\}\cdot(6k^2 + 10k)$ in which $H$ has treewidth at most $\binom{k+5}{4}-1$.
\end{cor}

>>>>

\begin{thm}
\label{kPlanarBasic}
Every $k$-planar graph is a subgraph of $H\boxtimes P$, for some graph $H$ of treewidth $(18k^2+48k+30)\binom{k+4}{3}-1$ and for some path $P$. 
\end{thm}

@@

\begin{thm}
\label{k-planar}
Every $k$-planar graph has an $H$-partition of layered width at most $18k^2 + 48k+30$ in which $H$ has treewidth at most $\binom{k+4}{3}-1$. %$\treewidth$.
\end{thm}

@@

\begin{thm}
\label{g-k-planar}
Every $(g,k)$-planar graph has an $H$-partition of layered width at most 
$\max\{2g,3\}\cdot(6k^2+16k+10)$ in which $H$ has treewidth at most $\binom{k+5}{4}-1$.
\end{thm}

@@

We can reduce the layered width of the $H$-partition of $G$ from $O(k^3)$ to $O(k^2)$ by observing that the dummy vertices in $V(G_0)\setminus V(G)$ do not contribute to the layered width of this partition.  In this setting, the proof of \cref{clm:general-width} is simpler since each vertex $w\in Y_x$ contributes at most two vertices to $V_i\cap Y_x$.  More precisely, each path $P\in\mathcal{P}$ containing an internal (dummy) vertex $w\in Y_x\cap (V_{i-j}\cup V_{i+j})$ contributes: (i)~at most two vertices to $S_x\cap V_i$ for $j\in\{0,\ldots,\floor{(k+1)/2}\}$; (ii)~at most one vertex to $S_x\cap V_j$ for $j\in\{\floor{(k+1)/2}+1,\ldots,k+1\}$; or (iii)~no vertices to $S_x\cap V_j$ for $j > k+1$.
Redoing the calculation at the end of the proof of \cref{clm:general-width} then yields 
\begin{align*}
    |S_x\cap Y_i| \le d\ell\left( 
       2 
       + 4\left\lfloor\frac{k+1}{2}\right\rfloor 
       + 2\left\lceil\frac{k+1}{2}\right\rceil
       \right)
       & =
       d\ell\left(
         2 + 2(k+1) + 2\left\lfloor\frac{k+1}{2}\right\rfloor
       \right) \\
       & \le
       d\ell(3k+5) 
      = 18k+30 \enspace .
\end{align*}    
With this change, the layered width of the partition given by \cref{modern-major-general} becomes $(18k+30)(k+1)=18k^2+48k+30$.  This establishes \cref{k-planar}.
% Note to self-and-others. In our original SODA submission we had 18k^2+30k here because we forgot that the layer compression step has compress k+1 consecutive layers, not just k consecutive layers.

@@

\begin{cor}\corlabel{first}
  For every $k$-planar graph $G$, 
  %$\pi(G)\le 12(2k+3)(k+1)\times 4^{(k+2)(k+3)(k+4)/6}$.
  $\pi(G)\le (18k^2+48k+30) 4^{\binom{k+4}{3}}$.
%  Every $k$-planar graph has a layered $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\binom{k+4}{3}}-1$.
\end{cor}

@@

\begin{cor}
  For every $(g,k)$-planar graph $G$, 
  \[ 
%    \pi(G)\le \max\{4g(3k+1),12(2k+3)(k+1)\}\times 4^{(k+2)(k+3)(k+4)/6} 
    \pi(G)\le \max\{2g,3\}\cdot(6k^2+16k+10) 4^{\binom{k+5}{4}}. 
     \]
%  \label{g-k-planar} Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most  
% $\max\{2g,3\}\cdot(6k^2 + 10k)$ in which $H$ has treewidth at most $\binom{k+5}{4}$.
\end{cor}

@@

\begin{cor}
\label{pck}
For every $k$-planar graph $G$, 
%for some integer $t\leq \binom{k+4}{3}-1$, 
%$$\chi_p(G)\le (18k^2+30k)(p+1) (t+1)\tbinom{p+t}{t}\enspace.$$
\[ \chi_p(G)\le (18k^2+48k+30)(p+1) \binom{k+4}{3} \binom{p+ \binom{k+4}{3}-1}{ \binom{k+4}{3}-1}\enspace. \]
%\note{DW}{I previously wrote ``for some integer $t\leq \binom{k+4}{3}-1$'' thinking that $\binom{p+t}{t}$ might not be monotone with $t$, but it is because of the $+t$.}
%Every $k$-planar graph has a layered $H$-partition of layered width at most $18k^2 + 30k$ in which $H$ has treewidth at most $\binom{k+4}{3}-1$. %$\treewidth$.
%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$. 
\end{cor}

@@

\begin{cor}
\label{pcgk}
For every $(g,k)$-planar graph $G$,  
%for some integer $t\leq \binom{k+5}{4}-1$, 
%$$\chi_p(G) \le \max\{2g,3\}\cdot(6k^2 + 10k) (p+1) (t+1)\tbinom{p+t}{t}\enspace.$$
\[ \chi_p(G) \le \max\{2g,3\}\cdot(6k^2+16k+10) (p+1) \binom{k+5}{4}\binom{p+\binom{k+5}{4}-1}{\binom{k+5}{4}-1}\enspace. \]
%O(gk^2)\times p^{O(k^3)}$. 
%%Every graph $G$ that has a layered $H$-partition of layered width at most $\ell$ in which $H$ has treewidth at most $t$ satisfies $\chi_p(G)\le \ell (p+1) (t+1)\binom{p+t}{t}$. 
%Every $(g,k)$-planar graph has a layered $H$-partition of layered width at most 
%$\max\{2g,3\}\cdot(6k^2 + 10k)$ in which $H$ has treewidth at most $\binom{k+5}{4}-1$.
\end{cor}
